---
title: "Event date data practice in R"
output: html_document
date: "2026-09-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is a markdown practice doc for some basic analysis in R working with data and event data. This is created in markdown for more effective commenting and highlighting questions for practice rather than usual R commenting. 


```{r setup, include=FALSE}

# load data
options(scipen=999) # turn off scientific numerical notation

`%notin%` <- Negate(`%in%`) # custom function for filtering data

# Specify packages
packages <-
  c("tidyverse","janitor","stringr","data.table",# general data wrangling
    "readxl","readr",
    "lubridate"
  ) 

# Install packages if not already installed
install.packages(setdiff(packages, rownames(installed.packages())))

# Load packages
sapply(packages, require, character.only = TRUE)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

data <- read_csv("food_delivery_scooter_events.csv") 

glimpse(data)
# unique list of data to view
unique_vals <- lapply(data, unique)


```


### Question 1: How long does each delivery take, and which stores are fastest or slowest on average?


```{r echo=FALSE}
time_taken_df <- data %>%
  filter(event_type %in% c("order_received","delivery_complete")) %>%
  group_by(delivery_id) %>%
  arrange(delivery_id, event_timestamp) %>%
  mutate(time_taken = event_timestamp - lag(event_timestamp))

time_taken_deliv <- time_taken_df %>%
  select(delivery_id, time_taken) %>%
  filter(!is.na(time_taken))

time_taken_deliv_all <- time_taken_deliv %>%
  ungroup() %>%
  summarise(time_taken_avg = mean(time_taken)) %>% pull()

```
```{r}

time_taken %>%
  group_by(store_name) %>%
  summarise(time_taken_avg = mean(time_taken, na.rm = TRUE)) %>%
  arrange(time_taken_avg)

```

- Time taken on average for all deliveries is `r round(time_taken_deliv_all,2)`. 
- 
  

