---
title: "Nordic trade with the UK"
output: html_document
date: "2026-06-05"
---



```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen=999) # turn off scientific numerical notation
# initial set up
# load in packages/API functions/data

library(tidyverse)
library(httr)
library(jsonlite)
library(furrr)
library(lubridate)
# Load API functions
source("api_functions.R")
source("data_processing.R")


```


## Overall UK-Nordic imports

In 2025 the UK imported £`r uk_totals[["2025"]]`M from around the world and £`r nordic_import_totals[["2025"]]`m from all Nordic countries combined. This is up in nominal terms since 2019 by £`r nordic_import_totals[["2025"]] - nordic_import_totals[["2019"]]`M. Overall Nordic share of UK imports is relatively small at `r nordic_import_share_uk.25`% in 2025 which is flat since 2019 however peaking in 2022 at `r nordic_import_share_uk.22`. Norway is the largest exporter into the UK of all Nordic countries, with £`r no_import_totals[["2025"]]`m in 2025, which is up `r no_import_totals[["2025"]] / no_import_totals[["2019"]] - 1`% since 2019 but also down X% form the 2022 peak of X% due to Russian Sanctions and importation of fuels.

## Share of Nordic imports. 

Table 1 summarises each Nordics nations overall goods imports with the UK in 2025. Norway is the largest share of all Nordic imports into the UK of `r no_nordic_share[["2025"]]`%, followed by Sweden (`r se_nordic_share[["2025"]]`%), Denmark (`r dk_nordic_share[["2025"]]`%), Finland (`r fi_nordic_share[["2025"]]`%) and then Iceland of (`r is_nordic_share[["2025"]]`%). These proportions while the scales have remained siilar have changed to varying degrees. Denmarks share has decreased by `r dk_nordic_share[["2025"]] - dk_nordic_share[["2019"]] ` percentage points while Norways has increased by `r no_nordic_share[["2025"]] - no_nordic_share[["2019"]]` percentage points since 2019. As Chart 1 shows Norways significance to Nordic imports with the UK has only increased since 2019 with an average share of `r round(mean(no_nordic_share),2)`%. 


#### Table 1: UK-Nordic imports in 2025
```{r}
library(knitr)
library(kableExtra)

uk_share_yr %>%
  filter(year == "2025") %>%
  arrange(-total_imports) %>%
  rename(
    Year = year,
    Country = country,
    `Total Imports (£m)` = total_imports
  ) %>%
  kable() %>%
  kable_styling() 

```



### Chart1: Share of overall UK-Nordic imports (%)

```{r, fig.width=12}


df <-  uk_share_yr %>% ungroup()

ordered_levels <- df %>%
  arrange(share_of_nordic) %>%
  pull(country) %>%
  unique()

df$country <- factor(df$country, levels = ordered_levels)




# stacked area plot for nordic share:

plot <- df %>%
  ggplot(
    aes(
      x=year, 
      y=share_of_nordic, 
      group = country,
      fill = country
     )
    )+
  geom_area(alpha = 0.5) + 
  theme_default +
  scale_fill_manual(
    #values = c( "#12436D" , "#0965A0" , "#2493D9" , "#7BC7F8" , "#777d96"),
    values = c( "#777d96" , "#7BC7F8" , "#2493D9" , "#0965A0" , "#12436D"),
    labels = c('Iceland', 'Finland','Denmark','Sweden','Norway')
  ) +
  labs(x="",  y="")+
  scale_y_continuous(labels = scales::percent)+
  theme(legend.title = element_blank())

plot

```


## Nordic time-series trend
In nominal terms, trade across Norway, Sweden and Iceland have increased while Denmark and Finland have both slightly decreased since 2019 as shown in Table 2. The average rate increase of X% per year across the block, with Norway the highest of X% and Finland the lowest with x%. 


```{r, fig.width=12}

uk_share_yr %>%
  mutate(year = as.integer(year)) %>%
  ggplot(aes(year, total_imports)) +
  geom_line(colour = "#1f78b4", linewidth = 1) +
  #facet_wrap(~ country, scales = "free_y") +
  scale_x_continuous(breaks = 2019:2025) +
  facet_wrap(~ country, scales = "free")+
  theme_default+
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.ticks.x = element_line(colour = "black", linewidth = 0.3),
    )


```


